---
title: "Test and Validation Error"
output: html_notebook
---

This script will test various statistical models against the various datasets to determine which would work best for the development of a TFR model.  This work is in support of the paper: “The Development of a TFR Gene Expression Data Set using Random Forest”
The work required an excel file containing the training and test dataset.  This can be set in the code.
 

These are the libraries used for this project:
```{r message=FALSE, include=FALSE}
library(grid)
library(corrplot)
library(caret)
library(e1071)
library(ROCR)
library(ggplot2)
library(GGally)
library(PerformanceAnalytics)
library(factoextra)
library(corrplot)
library(Rtsne)
library(FactoMineR)
library(ggplot2)
library(factoextra)
library(survminer)
library(ggcorrplot)
library(readr)
library(circlize)
library(readxl)
library(stringr)
library(reshape)
library(psych)
library(ComplexHeatmap)
library(ggpubr)
library(readr)
library(gridExtra)
library(cowplot)
library(MASS)
library(fitdistrplus)
#BiocManager::install("preprocessCore")
library(preprocessCore)
library(MASS)
library(class)
library(ggplot2)
library(reshape2)
library(ROCR)
library(e1071)
library(GGally)
library(klaR)
library(randomForest)
```

This is a function used to get the metrics from the perfomrance package. 
```{r}
summPreds <- function(inpPred,inpTruth,inpMetrNms=c("err","acc","sens","spec")) {
  retVals <- numeric()
  for ( metrTmp in inpMetrNms ) {
    retVals[metrTmp] <- performance(prediction(inpPred,inpTruth),measure=metrTmp)@y.values[[1]][2]
  }
  retVals
}
```


Loading in the training and test data:
```{r include=FALSE}
model = NULL
library(here)
here("/Data/Training_Test_Set.xlsx")
getwd()
library(readxl)
model <- read_excel("C:/Users/Alos Diallo/Documents/GitHub/TFR_Model/Data/Training_Test_Set.xlsx")

library(tidyverse)
model %>% select_if(negate(is.numeric))

```

Getting some information on the data.  You don't have to run this.
```{r}
dim(model)
size = ncol(model) -1
multi.hist(model[,2:size]) 
```

Here we are testing out the different models.

```{r testerr,warning=FALSE}

dfTmpLR <- NULL
dfTmpRF <- NULL
dfTmpNB <- NULL
dfTmpLDA <- NULL
dfTmpSVM <- NULL
dfTmp = NULL
svmTuneRes = NULL
Train = NULL
Test = NULL
dfTmpKNN = NULL

for ( iResample in 1:2 ) {
  for ( iSim in 1:100 ) {
    bTrain = sample(1:nrow(model),round(0.7*nrow(model)))
    Train <- model[bTrain,]
    Test <- model[-bTrain,]
    Test = as.data.frame(Test)
    Train$cat = as.numeric(as.factor(Train$cat))
    Test$cat = as.numeric(as.factor(Test$cat))
    if ( iResample == 2 ) {
      bTrain = sample(1:nrow(model),round(0.7*nrow(model)))
      Train <- model[bTrain,]
      Test <- model[-bTrain,]
      Test = as.data.frame(Test)
      Train$cat = as.numeric(as.factor(Train$cat))
      Test$cat = as.numeric(as.factor(Test$cat))
    }
# logistic regression:
glmTrain <- glm(as.factor(Train$cat)~.,data=Train[,2:size],family=binomial)
glmTestPred <- predict(glmTrain, newdata=Test[,2:size], type="response") > 0.5

tmpVals <- summPreds(as.numeric(glmTestPred)+1,as.factor(Test$cat))
dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="LR",metric=names(tmpVals),value=tmpVals))
dfTmpLR <- rbind(dfTmpLR,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="LR",metric=names(tmpVals),value=tmpVals))
  
#RF
Y = factor(as.numeric(as.factor(Train$cat)))
rfTmp <- randomForest(Y~.,data=Train[,2:size],mTry = 10)
rfTestPred <- predict(rfTmp,newdata=Test)
mseTest <- mean((as.numeric(factor(Test$cat))-as.numeric(rfTestPred))) #Calculating the mean squared error

tmpVals <- summPreds(as.numeric(factor(Test$cat)),as.numeric(rfTestPred))
dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="RF",metric=names(tmpVals),value=tmpVals))
dfTmpRF <- rbind(dfTmpRF,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="RF",metric=names(tmpVals),value=tmpVals))
# LDA:
ldaTrain <- lda(Train$cat~.,data=Train[,2:size])
ldaTestPred <- predict(ldaTrain, newdata=Test[,2:size])

tmpVals <- summPreds(as.numeric(ldaTestPred$class),as.factor(Test$cat))
dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="LDA",metric=names(tmpVals),value=tmpVals))
dfTmpLDA <- rbind(dfTmpLDA,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="LDA",metric=names(tmpVals),value=tmpVals))
# NB:
Y = factor(as.numeric(as.factor(Train$cat)))
nbTrain <- naiveBayes(Y~.,data=Train[,2:size])
nbTestPred = predict(nbTrain, newdata=Test[,2:size])

tmpVals = summPreds(as.numeric(nbTestPred),as.numeric(as.factor(Test$cat)))
dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="NB",metric=names(tmpVals),value=tmpVals))
dfTmpNB <- rbind(dfTmpNB,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="NB",metric=names(tmpVals),value=tmpVals))
#SVM
svmResult = svm(Train$cat~.,data=Train[,2:size],kernel="radial",cost=20,gamma = .1)
svmTestPred = predict(svmResult, newdata=Test[,2:size])

tmpVals = summPreds(predict(svmResult,newdata=Test[,2:size]),Test$cat)
dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="SVM",metric=names(tmpVals),value=tmpVals))
dfTmpSVM <- rbind(dfTmpSVM,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="SVM",metric=names(tmpVals),value=tmpVals))
# KNN:
for ( kTmp in c(10,20,150,300) ) {
  knnTestPred <- knn(Train[,2:size],Test[,2:size],Train$cat,k=kTmp)
  knnTrainPred <- knn(Train[,2:size],Train[,2:size],Train$cat,k=kTmp)

  tmpVals <- summPreds(as.numeric(knnTestPred),as.numeric(Test$cat))
  
  dfTmp <- rbind(dfTmp,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="KNN",metric=names(tmpVals),value=tmpVals))
  dfTmpKNN <- rbind(dfTmpKNN,data.frame(resample=c("Test Error BS1","Test Error BS2")[iResample],type="KNN",metric=names(tmpVals),value=tmpVals))
 
  }
  }
}
```

Here we have the performance metrics for each model.
```{r}
p = NULL
p = ggplot(dfTmp,aes(x=type,y=100*value,colour = type)) + geom_boxplot(fill="white") + geom_point() + facet_wrap(~resample+metric,ncol=4,scales="free") + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme_bw()

p
```


