---
title: "QC for tfr project"
output: html_notebook
---

Loading libraries
```{r message=FALSE, include=FALSE}
library(neuralnet)
library(readxl)
library(grid)
library(corrplot)
library(caret)
library(e1071)
library(ROCR)
library(ggplot2)
library(GGally)
library(PerformanceAnalytics)
library(factoextra)
library(corrplot)
library(Rtsne)
library(FactoMineR)
library(ggplot2)
library(factoextra)
library(survminer)
library(ggcorrplot)
library(readr)
library(circlize)
library(readxl)
library(stringr)
library(reshape)
library(psych)
library(ComplexHeatmap)
library(ggpubr)
library(readr)
library(gridExtra)
library(cowplot)
library(MASS)
library(fitdistrplus)
#BiocManager::install("preprocessCore")
library(preprocessCore)
library(MASS)
library(class)
library(ggplot2)
library(reshape2)
library(ROCR)
library(e1071)
library(GGally)
library(klaR)
library(randomForest)
library(tidyverse)
library("Rgraphviz")
library(CePa)
library(corrplot)
library(svglite)
library('biomaRt')
library("EnsDb.Hsapiens.v79")
library( "dplyr" )
```


```{r}
model_4_21_95wRAW = NULL
model_2_7_21_new = NULL
library(readxl)
model_2_7_21_new <- read_excel("F:/Research/Sage/2021 Model/model_2_7_21_newV4.xlsx")

#model_2_7_21_new2 = apply(model_2_7_21_new[,2:104],2,as.numeric)
#model_4_21_95wRAW <- cbind(model_2_7_21_new[,1],model_2_7_21_new2)

#View(model_4_21_95wRAW)  
model_4_21_95wRAW <- model_2_7_21_new

library(tidyverse)
model_4_21_95wRAW %>% select_if(negate(is.numeric))
```


```{r}
summPreds <- function(inpPred,inpTruth,inpMetrNms=c("err","acc","sens","spec")) {
  retVals <- numeric()
  for ( metrTmp in inpMetrNms ) {
    retVals[metrTmp] <- performance(prediction(inpPred,inpTruth),measure=metrTmp)@y.values[[1]][2]
  }
  retVals
}
```


```{r}

svg("hist.svg",    width=150, height=150,pointsize = 30)
multi.hist(model_4_21_95wRAW[,2:86]) 
dev.off()


dim(model_4_21_95wRAW)
multi.hist(model_4_21_95wRAW[,2:86]) 
#pairs(model_4_21_95wRAW[,2:86])
```


PCA
```{r}



#here we are computing PCA

prcompTmp <- prcomp(model_4_21_95wRAW[,2:86],center = TRUE,scale = TRUE)

#Here we have a plot of the variance by principle componant
plot(prcompTmp)

#This is just a fancy way of doing the same thing
fviz_eig(prcompTmp, addlabels=TRUE, ylim=c(0,60), geom = c("bar", "line"), barfill = "gold", barcolor="grey",linecolor = "red", ncp=10)+
labs(title = "Variance Plot ",
         x = "Principal Components", y = "% of variances")


all_var <- get_pca_var(prcompTmp)
all_var

#Correlation between variables and PCA
#It shows the importance of a principal component for a given observation (vector of original variables). You can go through the following link for details.
corrplot(all_var$cos2, is.corr=FALSE)

#To highlight the most contributing variables for each components
corrplot(all_var$contrib, is.corr=FALSE)    
# 
pdf("cor_plots.pdf")
corrplot(all_var$contrib, is.corr=FALSE) 
dev.off()

fviz_pca_var(prcompTmp, col.var = "black")


fviz_pca_var(prcompTmp, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
             )




```


QC for rows:
```{r}
 Train = NULL
    Test = NULL
    bTrain = sample(1:nrow(model_4_21_95wRAW),round(0.7*nrow(model_4_21_95wRAW)))
    Train <- model_4_21_95wRAW[bTrain,]
    Test <- model_4_21_95wRAW[-bTrain,]
    Test = as.data.frame(Test)
    Train$cat = as.numeric(as.factor(Train$cat))
    Test$cat = as.numeric(as.factor(Test$cat))
          
scrambled_outcome = sample(Train$cat)
scrambled_outcome_test = sample(Test$cat)
length(scrambled_outcome)
length(Train$cat)
```


```{r}

Y = factor(as.numeric(as.factor(scrambled_outcome)))
Y_test = factor(as.numeric(as.factor(scrambled_outcome_test)))
rfTmp <- randomForest(Y~.,data=Train[,2:86],mTry = 10)
rfTestPred <- predict(rfTmp,newdata=Test)
mseTest_scrambled <- mean((as.numeric(factor(Y_test))-as.numeric(rfTestPred))) #Calculating the mean squared error
tmpVals_scrambled <- summPreds(as.numeric(factor(Y_test)),as.numeric(rfTestPred))


```

```{r}
Y_s = factor(as.numeric(as.factor(scrambled_outcome)))
Y_test = factor(as.numeric(as.factor(scrambled_outcome_test)))
Y = factor(as.numeric(as.factor(Train$cat)))

dfTmp <- NULL
for ( iResample in 1:2 ) {
  for ( iSim in 1:100 ) {
 
  
  scrambled_outcome_eachTime = sample(Train$cat)
  scrambled_outcome_test_eachTime = sample(Test$cat)  
  Y_s_et = factor(as.numeric(as.factor(scrambled_outcome_eachTime)))
  Y_test_et = factor(as.numeric(as.factor(scrambled_outcome_test_eachTime)))  
    
    
  rfTmp <- randomForest(Y~.,data=Train[,2:86],mTry = 10)
  rfTestPred <- predict(rfTmp,newdata=Test[,2:86])
  mseTest <- mean((as.numeric(factor(Test$cat))-as.numeric(rfTestPred))) #Calculating the mean squared error
  tmpVals <- summPreds(as.numeric(factor(Test$cat)),as.numeric(rfTestPred))
  dfTmp <- rbind(dfTmp,data.frame(resample=c("Bootstrap 1","Bootstrap 2")[iResample],type="RF",metric=names(tmpVals),value=tmpVals))  
    
  rfTmpS <- randomForest(Y_s~.,data=Train[,2:86],mTry = 10)
  rfTestPredS <- predict(rfTmpS,newdata=Test[,2:86])
  mseTest_scrambled <- mean((as.numeric(factor(Y_test))-as.numeric(rfTestPredS))) #Calculating the mean squared error
  tmpVals_scrambled <- summPreds(as.numeric(factor(Y_test)),as.numeric(rfTestPredS))
  dfTmp <- rbind(dfTmp,data.frame(resample=c("Bootstrap 1","Bootstrap 2")[iResample],type="RF Scrambled",metric=names(tmpVals_scrambled),value=tmpVals_scrambled))

  rfTmpET <- randomForest(Y_s_et~.,data=Train[,2:86],mTry = 10)
  rfTestPredET <- predict(rfTmpET,newdata=Test[,2:86])
  mseTest_scrambledET <- mean((as.numeric(factor(Y_test_et))-as.numeric(rfTestPredET))) #Calculating the mean squared error
  tmpVals_scrambledET <- summPreds(as.numeric(factor(Y_test_et)),as.numeric(rfTestPredET))
  dfTmp <- rbind(dfTmp,data.frame(resample=c("Bootstrap 1","Bootstrap 2")[iResample],type="RF Scrambled Each Time",metric=names(tmpVals_scrambledET),value=tmpVals_scrambledET))

  }
}
```

```{r}
library(RColorBrewer)

p = ggplot(dfTmp,aes(x=type,y=100*value,colour = type)) + geom_boxplot(fill="white") + geom_point() + facet_wrap(~resample+metric,ncol=4,scales="free") + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme_bw() + scale_color_brewer(palette = "Dark2")

p
```


Measuring how correlated the columns are with respect to N and Y.  Here I am looking at for a given column how correlated the outcomes are for each column then building a histogram showing this. 
```{r}
library(corrr)
#model_4_21_95wRAW[,2:103]


Corrs = NULL
maxS = ncol(model_4_21_95wRAW) -1
for(i in 2:maxS){
  x = model_4_21_95wRAW[2:1023,i]
  y = model_4_21_95wRAW[1024:2077,i]
  x = sample(as.numeric(unlist(x)),1000)
  y = sample(as.numeric(unlist(y)),1000)
  Corrs[i] = cor(x,y)
}  
  
hist(Corrs, main = "Histogram of Correlation for Y and N",xlab =  "Pearson Correlation Values") 
```
Here I am looking at the correlation between rows. In this case the Y vs the N
```{r}

y_s = model_4_21_95wRAW[2:1023,]
n_s = model_4_21_95wRAW[1024:2077,]

yT = t(y_s)
nT = t(n_s)

dim(yT)
dim(nT)
nT = nT[2:86,]
nT = as.numeric(nT)
nT = matrix(nT,86,1054)
nT = nT[,1:1022]
yT = yT[2:86,]
yT = as.numeric(yT)
yT = matrix(yT,86,1022)

No_cor = cor(nT)

Yes_cor = cor(yT)

total_cor = cor(nT,yT)



Yes_D = density(Yes_cor)
No_D = density(No_cor,na.omit)
Total_D = density(total_cor)


NewBlue = rgb(0,0,1,0.8)
NewGreen = rgb(0,1,0,0.6)
NewRed = rgb(1,0,0,0.6)


pdf("my_plot.pdf")
heatmap(No_cor,symm = T,main = "Heatmap: Correlation of No")
heatmap(Yes_cor,symm = T,main = "Heatmap: Correlation of Yes")
heatmap(total_cor,symm = T,main = "Heatmap: Correlation of No vs Yes")
plot(Total_D)
polygon(Total_D,col = NewRed)
polygon(No_D,col = NewBlue)
polygon(Yes_D,col = NewGreen )

legend("topright", inset=.02, title="Comparing correlation",
   c("No","Yes","No vs Yes"), fill= c(NewBlue,NewGreen,NewRed),horiz=FALSE, cex=0.8)
dev.off()
```

Testing external data to see how it compares.
```{r}

gtex = read.gct("F:/Research/Sage/QC_dataSet/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct")

```

```{r}
dim(gtex)
```


Making a subset of the gtex based on my list of genes
```{r}
Ys = model_4_21_95wRAW$Feature.ID
Ys = as.character(Ys[2:1023])
Ns = model_4_21_95wRAW$Feature.ID
Ns = as.character(Ns[1024:2077])




#gene_list = as.character(unlist(gene_list))
#small_list_of_genes <- subset(df_res, rownames(df_res) %in% gene_list)
```

converting the symboles to gene names
```{r}
gtex2 = gtex
GtexNames = rownames(gtex2)
GtexNames2 <- gsub(GtexNames, pattern="\\.[0-9]+$", replacement="")

GtexNames = as.character(unlist(GtexNames2)) 
#geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys = GtexNames , keytype = "GENEID", columns = c("SYMBOL","GENEID"))


v = c("ENSG00000263529","ENSG00000273274","ENSG00000160062","ENSG00000176261","ENSG00000200591","ENSG00000162521","ENSG00000162520","ENSG00000224409") 

mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
G_list <- getBM(filters = "ensembl_gene_id", 
                attributes = c("ensembl_gene_id","hgnc_symbol"),
                values = GtexNames, mart = mart)
print(G_list)

#Here I'm just trying to get some random gene names for the random gene lists.
randomnames = sample(G_list$hgnc_symbol,3000)
write.csv(randomnames, file = "random_names.csv")

```

Now I am taking the converted names and subsetting the gtex data with them
```{r}
Ys = lapply(Ys, toupper)
unlist(Ys)
Ns = lapply(Ns, toupper)
unlist(Ns)
small_list_of_genes_Ys <- subset(G_list, G_list[,2] %in% Ys)

small_list_of_genes_Ns <- subset(G_list, G_list[,2] %in% Ns)

rownames(gtex2) = GtexNames2
small_list_of_genes_Ys_G <- subset(gtex2, rownames(gtex2) %in% small_list_of_genes_Ys[,1])

small_list_of_genes_Ns_G <- subset(gtex2, rownames(gtex2) %in% small_list_of_genes_Ns[,1])
```

Now I want to generate correlations
```{r}
small_list_of_genes_Ys_G
small_list_of_genes_Ns_G

bTrain <- sample(rep(c(TRUE,FALSE),length.out=ncol(small_list_of_genes_Ys_G)),replace = FALSE,size = 1000) 

small_list_of_genes_Y <- small_list_of_genes_Ys_G[,bTrain]
small_list_of_genes_N <- small_list_of_genes_Ns_G[,bTrain] 
smallyT = t(small_list_of_genes_Y)
smallnT = t(small_list_of_genes_N)

dim(smallyT)
smallyT = smallyT[1:600,1:600]
dim(smallyT)
dim(smallnT)
smallnT = smallnT[1:600,1:600]
dim(smallnT)

rownames(smallyT) = NULL
colnames(smallyT) = NULL
rownames(smallnT) = NULL
colnames(smallnT) = NULL


smallnT = as.numeric(smallnT)
smallnT = matrix(smallnT,600,600)
smallyT = as.numeric(smallyT)
smallyT = matrix(smallyT,600,600)

No_cor = cor(smallnT)
Yes_cor = cor(smallyT)

total_cor = cor(smallnT,smallyT)



Yes_D = density(Yes_cor)
No_D = density(No_cor)
Total_D = density(total_cor)


NewBlue = rgb(0,0,1,0.8)
NewGreen = rgb(0,1,0,0.6)
NewRed = rgb(1,0,0,0.6)


pdf("GTEX_my_plot.pdf")
heatmap(No_cor,symm = T,main = "Heatmap: Correlation of No")
heatmap(Yes_cor,symm = T,main = "Heatmap: Correlation of Yes")
heatmap(total_cor,symm = T,main = "Heatmap: Correlation of No vs Yes")
plot(No_D)
polygon(No_D,col = NewBlue)
polygon(Yes_D,col = NewGreen )
polygon(Total_D,col = NewRed)
legend("topright", inset=.02, title="Comparing correlation GTEX",
   c("No","Yes","No vs Yes"), fill= c(NewBlue,NewGreen,NewRed),horiz=FALSE, cex=0.8)
dev.off()
```

```{r}

small_list_of_genes_Y <- small_list_of_genes_Ys_G[,bTrain]
small_list_of_genes_N <- small_list_of_genes_Ns_G[,bTrain] 
addYs = rep('Y',time = nrow(small_list_of_genes_Y))
small_list_of_genes_Y = cbind(small_list_of_genes_Y,addYs)
addNs = rep('N',time = nrow(small_list_of_genes_N))
small_list_of_genes_N = cbind(small_list_of_genes_N,addNs)
data = rbind(small_list_of_genes_Y,small_list_of_genes_N)
```


```{r}
summPreds <- function(inpPred,inpTruth,inpMetrNms=c("err","acc","sens","spec")) {
  retVals <- numeric()
  for ( metrTmp in inpMetrNms ) {
    retVals[metrTmp] <- performance(prediction(inpPred,inpTruth),measure=metrTmp)@y.values[[1]][2]
  }
  retVals
}
```


```{r}
#Make sure to re-run bTrain

dfTmp <- NULL
for ( iResample in 1:2 ) {
  for ( iSim in 1:1500 ) {
    
    #make this into a loop tomorrow.
    colsToUse = sample(2:1000,40)
    bTrain = sample(1:nrow(data),round(0.7*nrow(data)))
    Train <- data[bTrain,]
    Test <- data[-bTrain,]
    Y = factor(as.numeric(as.factor(Train[,8774])))
    
    rfTmp = NULL
    rfTmp <- randomForest(Y~.,data=Train[,colsToUse],mTry = 10)
    Y_test = factor(as.numeric(as.factor(Test[,8774])))
    rfTestPred <- predict(rfTmp,newdata=Test[,colsToUse])
    mseTest <- mean((as.numeric(factor(Y_test))-as.numeric(rfTestPred))) 
    tmpVals <- summPreds(as.numeric(factor(Y_test)),as.numeric(rfTestPred))  
    dfTmp <- rbind(dfTmp,data.frame(resample=c("Bootstrap Round 1","Bootstrap Round 2")[iResample],type="RF GTEX",metric=names(tmpVals),value=tmpVals))
    
  }
}
```


```{r}
p = ggplot(dfTmp,aes(x=type,y=100*value,colour = type)) + geom_boxplot(fill="white") + geom_point() + facet_wrap(~resample+metric,ncol=4,scales="free") + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme_bw()

p
```


Looking at the correlation of the columns
```{r}
library(corrr)
library( "dplyr" )
#model_4_21_95wRAW[,2:103]
Yes_columns = NULL
No_columns = NULL
Yes_columns = dplyr::filter( model_4_21_95wRAW, cat == "y" )
Yes_columns = Yes_columns[,1:103]
Yes_columns = as.numeric(unlist(Yes_columns))
Yes_columns = matrix(Yes_columns,1022,103)
Ycorrelation = cor(Yes_columns)
No_columns = dplyr::filter( model_4_21_95wRAW, cat == "n" )
No_columns = No_columns[,1:103]
No_columns = as.numeric(unlist(No_columns))
No_columns = matrix(No_columns,1055,103)
Ncorrelation = cor(No_columns)

corrplot(Ycorrelation) 
corrplot(Ncorrelation) 
total_cor = cor(Yes_columns,No_columns[1:1022,])

Yes_D = density(Ycorrelation[2:103,2:103])
No_D = density(Ncorrelation[2:103,2:103])
Total_D = density(total_cor[2:103,2:103])


NewBlue = rgb(0,0,1,0.8)
NewGreen = rgb(0,1,0,0.6)
NewRed = rgb(1,0,0,0.6)


pdf("COl_plot.pdf")
heatmap(No_cor,symm = T,main = "Heatmap: Correlation of No")
heatmap(Yes_cor,symm = T,main = "Heatmap: Correlation of Yes")
#heatmap(total_cor,symm = T,main = "Heatmap: Correlation of No vs Yes")
plot(No_D)
polygon(No_D,col = NewBlue)
polygon(Yes_D,col = NewGreen )
polygon(Total_D,col = NewRed)
legend("topright", inset=.02, title="Comparing correlation for Columns",
   c("No","Yes","Yes vs No"), fill= c(NewBlue,NewGreen,NewRed),horiz=FALSE, cex=0.8)
#legend("topright", inset=.02, title="Comparing correlation for Columns",
#   c("No","Yes"), fill= c(NewBlue,NewGreen,NewRed),horiz=FALSE, cex=0.8)
dev.off()



```